CreateFiguresReview

### 2025-06-14 manuscript/ResponseReview1.Rmd

Response: This is a tutorial specifically focused on Bayesian methods and cognitive experimental psychology data, such as response times and saccade latencies.
Analyzing such event latencies with EHA can give better answers about the time course of psychological processes compared to (a) ANOVA on means and (b) currently popular quantile-based distributional methods. 
Information about the time course of cognitive processes is present in the detailed shape of the hazard function which quantifies the moment-to-moment changes in behavioral performance, in terms of the risk of response occurrence, and the accuracy of responses emitted. EHA thus allows answering research questions about the onset of psychological effects, about time-locking of effects to stimuli, and about how the effect changes on multiple time scales, i.e., with the passage of time within trials, across trials, across blocks within a session, etc. 
Applying EHA also has implications for experimental design, such as response deadlines, timer on trial onset, etc.

> On page 7, it states that there are not tutorials aimed at psychological time-to-event data. Yet, there are at least a few existing empirical article examples with accompanying tutorials in psychology (listed below), and arguably some of the popular textbooks are also geared toward psychological scientists (e.g., Singer & Willett). Similarly, many of the textbooks contain code to accompany their text. Regardless, even if those are not counted, I still think it will be beneficial for readers if the authors can make a clear case for what is unique about psychological time-to-event data that necessitates a stand alone tutorial/ what gaps in the existing textbooks/empirical articles/tutorials the present tutorial serves to fill.

Response: We believe a stand-alone tutorial for cognitive experimental psychology is necessary because (a) psychologists are reluctant to abandon ANOVA and switch to EHA/SAT for several reasons, (b) as far as we know, none of the existing tutorials focuses on Bayesian methods.

> Lougheed, J. P., Benson, L., Cole, P. M., & Ram, N. (2019). Multilevel survival analysis: Studying the timing of children’s recurring behaviors. Developmental Psychology, 55(1), 53.

> Elmer, T., van Duijn, M. A., Ram, N., & Bringmann, L. F. (2023). Modeling categorical time-to-event data: The example of social interaction dynamics captured with event-contingent experience sampling methods. Psychological Methods.

> Mills, M. (2011). The fundamentals of survival and event history analysis. Introducing Survival Analysis and Event History Analysis. London: SAGE Publications, 1-17.

Response: We included these references in the revised ms.

> <b>Organization of the tutorial:</b>The current version of the tutorial is much longer than a typical AMPPStutorial. There are a few possibilities for dealing with this. First, it may be that the tutorial could be split into 2, such as splitting into tutorials 1-3 and tutorial 4 on its own. Another possibility is that the focus could be primarily on the Bayesian models, and then leave the frequentist models for either a supplement, or to only live in the RMarkdown. This possibility may also be beneficial due to the model convergence issues with at least one of the frequentist models (which is common with multilevel data and why many psychological scientists are moving to the Bayesian framework). A third possibility is that the overall tutorial text could be greatly streamlined, with major cuts to both the introduction and the discussion sections.

Response: We focus primarily on the Bayesian models in the revised ms.  We also made major cuts to both the introduction and the discussion sections.

> <b>Clarity and accuracy: </b>
-There are several instances where jargon terms and key concepts pertinent to survival analysis need to be defined and/or described in slightly more detail than what is currently provided. For example, clearer descriptions of censoring and risk sets (plus what it means to be at risk) should be given in the introduction.

Response: We provide clear descriptions of censoring and risk sets on page x of the revised ms.

> Additionally, many key details, such as the EHA equations are relegated to supplementary materials. In my opinion, any key information that makes the tutorial more accessible and precise should not be in supplementary materials.

Response: We placed the equations for h(t), S(t), and ca(t) in the revised ms. on page x. The equations for P(t) are still in the supplementary material.

> There are many abbreviations that were difficult to follow. EHA is arguably standard, but SAT (for speed/accuracy trade-off) is less familiar, and may therefore benefit from not being an abbreviation. Similarly referring to hazard function as h(t), conditional accuracy function as c(t) was less clear than just saying hazard function and conditional accuracy function.

Response: Although SAT is rather standard in cognitive experimental psychology, we tried to limit the use of abbreviations in main body of the revised ms.

> It would be helpful if the research questions that can be addressed by the methods presented are clearly presented in each section. I think that currently they are getting lost in the tutorial.

Response:

> There are several opportunities to describe the method more precisely. For example, the discrete time hazard function gives a ‘conditional’ probability (pg. 9).

Response:

> Additionally, it may be helpful to use more standard notations such as h(t_ij) to indicate the conditional probability that the event occurs in bin j for individual i, given that the individual had not already experienced the event in a previous time period (lines 159-160, pg. 9).

Response:

>  Similarly, the survival probability is the probability that the event occurs after bin j, *given that the individual lhas survived up until bin j*.

Response:

> Throughout the tutorial, reference to individuals in the notation is missing.

Response: We added reference to individuals in the notation.

> Another inaccuracy is that in the discrete time framework, it is not an instantaneous likelihood (line 161, pg. 9).

Response: Indeed, only continuous-time hazard can be called an instantaneous likelihood.

>  <b>Introduction:</b>
-Section 1.1 – I had a difficult time understanding this section as the opening to the tutorial, and the last paragraph seemed out of place/abrupt. felt too vague to be understandable as the opening to the tutorial.

Response:

> Given that the speed-accuracy tradeoff analysis is described as central to this tutorial, it may benefit from being introduced earlier than pg. 7.

Response: We introduce SAT analysis already on page x in the revised ms.

> It seems that comparing and contrasting Bayesian and frequentist methods may be outside the scope of a tutorial (pg. 7)

Response: We agree and focus now on the Bayesian models in the revised ms.

> As mentioned above, tutorial 4 – while related – could potentially be its own stand-alone tutorial.

Response: We decided to keep tutorial 4.

> Censoring is a key component of survival analysis that is missing from the introduction

Response:

> It might be useful to indicate early on the distinctions between multilevel data, single events, and recurring events – and that this particular tutorial is only focused on single events.

Response: We included the following text on page x of the revised ms.: This tutorial deals with single-event analyses which focus on the time until a single event occurs in a trial (e.g., a button press), with repeated measures obtained for each individual. Recurrent event analysis deals with multiple events that can happen repeatedly to the same individual within a single observation period.

> It’s not clear why a discrete time framework is used given the granular temporal resolution of reaction time data – this would also mean that the data would not need to be separated into bins. Most textbooks (e.g., Allison’s) often indicate discrete time methods are common for temporal scales such as months, years, or decades (not milliseconds). Clearer justification for the discrete time framework and separating the data into bins would be helpful.

Response:

> Clearer description of the data would be helpful for readers (as well as only having this description in one place in the tutorial to be more concise). For example, on pg. 10 it is indicated that 200 trials of 1 experimental condition are simulated. Can the event occur at any time in those 200 trials? If the event occurs in trial 150, do the remaining 50 trials still occur, or does the time series end for that person? How are those extra trials after an event occurred handled in the data wrangling and analysis. Can events occur more than once?

Response:

> Section 2.2 (pg. 10, lines 168-169), if describing what ‘statisticians and mathematical psychologists say’, it would be helpful to provide supporting references.

Response: We provide supporting references on page x of the revised ms.

> The numbered list on pg. 11 contained jargon statements without explanations. Suggest to simplify and keep the focus on the data you are using and specific research questions you are using survival analysis as a tool to address.

Response: We removed the numbered list in the revised ms.

> In section 2.3, suggest to lead with the data and research questions of interest and then show how features of survival analysis make it a good method to examine the research questions. I also felt that up until now, the tutorial has been very confusing and hard to follow. It could be that starting the tutorial at section 2.3 (with a few extra details filled in), may be an alternative.

Response: We follow the suggestion of this reviewer to start the tutorial at section 2.3. That is, we immediately provide an example experiment/data set with a descriptive analysis that illustrates the benefits of applying EHA to psychological time-to-event data such as response times and saccade latencies. This new example data set is taken from Panis & Schmidt (2022).

> Again, section 2.3.1 (the majority of pg. 12) was unclear and difficult to follow for someone who already has knowledge of survival analysis, but not much familiarity with cognitive process/reaction time data.

Response:

> From what is written, I did not have a clear sense of how temporal states of cognitive processes and theory development could occur (pg. 14, lines 231-29). These ideas and the research questions are scattered in many places throughout the tutorial, and are not very clear/straightforward. It may be helpful to locate them in just one place in the tutorial and provide a clearer description to guide readers. Similarly, the statement on pg. 14 (lines 238-239) felt rather strong without any references, and also out of scope for a tutorial.

Response:

> It wasn’t clear how section 3.1 added value to the tutorial

Response:

> Section 3.2 – it would be helpful if there was clarification in the text on when it would not be possible to include varying intercepts and slopes (i.e., at least clarifying that random intercepts are a requirement, but sometimes random slopes are removed based on model convergence issues – however, this also begs the question of whether a model should even be presented if it does not converge when it includes random slopes…)

Response:

> Additionally in section 3.2, the text on pg. 16 is vague and does not give much detail to readers.

Response:

> Overall, the introduction spans across 16 pages, which feels long even for a standard empirical article, let alone a tutorial. As currently written, I think many readers will get lost before even getting to the tutorials section.

Response: We restructured and shortened the introduction.

> <b>Tutorials section:</b>
-Rather than only using h(t), and ca(t), suggest to use hazard function and conditional accuracy function along with their mathematical notation. Without having presented any equations, using terms such as h(t) and ca(t) seems odd.

Response: We try to limit the number of abbreviations used in the main body of the revised ms.

> Similarly, P(t) is never defined for readers.

Response: P(t)  is defined only in the supplementary material.

> Section 4.1.1. – it would be helpful to clarify that the risk set also pertains to individuals (not just bins)

Response: We clarify this on page x of the revised ms.

> The connection between trials and bins could be made clearer (this is in reference to the text on pg. 18, but I think the clarity would need to come earlier on in the tutorial when describing the data)

Response:

> Comprehension of the text on pg. 18 would be enhanced if both right and left censoring had been clearly introduced previously. Although left censoring is probably unlikely in the reaction time data used in the tutorial – if it is the case that this tutorial is meant for broader data types in psychology, left censoring would certainly be relevant to many other use cases.

Response: We introduce types of censoring on page x of the revised ms.

> Section 4.1.2 was a clearer data description than had been presented up until that point in the tutorial. Again, it is suggested that the manuscript/tutorial could be dramatically reduced in length if information was more concisely but descriptively written in a single place (e.g., data descriptions).

Response: We provide the information on data descriptions in a single place (page x of the revised ms.).

> Suggest to be more precise with the language on pg. 21 (line 355) – which data are nested within participants? The trial data? The binned data?

Response:

> pg. 22 (line 358) the way that censoring is described here is confusing.

Response:

> pg. 22 discusses warning messages and lets readers know these can be ignored but does not give details as to why. Since this is a didactic tutorial, it may be helpful to have at least some explanation as to why the warning can be ignored.

Response:

> pg. 22 – since this is a tutorial, it might also be helpful to unpack in more detail the section about identifying individuals who may be guessing etc – especially since these are very important considerations.

Response: We unpack in more detail how individuals can be identified that may be guessing, or that do not perform the task as intended.

> section 4.3.1 – (line 451) it would be helpful if more details is given on how to set the analysis time window. What counts as “enough data”? Given that this is a tutorial, these types of detail are critical for readers to be able to implement the method in their own work

Response:

> section 4.3.1 (line 453) – more detail is needed on the cloglog link function and how it differs from the more commonly known about logit link.

Response:

> section 4.3.1 (line 456) – what goes into the decision about whether to treat time as a categorical or continuous predictor?

Response:

> section 4.3.3. – it might be helpful to give a brief indicator up front to readers about the rationale/purpose of fitting both M0i and M1i, especially given that no interpretation is given for M0i. Similarly, it would be helpful to give a bit more detail on the purpose of the comparison of M0i and M1i (given that this is a didactic tutorial).

Response: We provide a brief description of the rationale of comparing models M0i and M1i on page x of the revised ms.

> section 4.3.6. – same thing here – since this is a tutorial, at least brief details on what the purpose of this section is would be helpful. It might also be helpful here to define what is meant by credible interval so that more novice readers don’t confuse it with a confidence interval.

Response:

> The term SOA is used on pg. 34, but had never previously been defined or described.

Response:

> The text on pg. 35 (lines 567-574) seems too in-depth for a tutorial

Response:

> pg. 36 – it would be helpful to give more detail on how to interpret the logit-ca scale

Response:

> section 4.7 – as previously mentioned, it is possible that tutorial 4 could be a stand alone tutorial. If it’s not though, more details are needed up front to tell readers what the purpose is and why this is useful.

Response:

> <b>Discussion section:</b> The Discussion section would benefit from being shortened (AMPPS guidelines indicate discussions should be a brief summary of their contents rather than a general discussion).

Response: We shortened the discussion section.

> is ‘how long’ a component of survival analysis – or is it more accurate to say ‘whether and when psychological states occur’?

Response: It is more accurate to say ‘whether and when’ psychological states occur. However, based on those results, we can deduce information about how long a psychological state lasts within a trial.

> pg. 47 (lines 753-761) – I had a difficult time understanding this section

Response: We rewrote this section.

> section 5.5 – unclear that the multi-state model is actually an example where a multi-state model would be relevant… (the conceptual example given seems like it would be more linked to either a hidden Markov model or recurring event survival model)

Response: We provide a better example in the revised ms.

>  <b>Figures:</b>
-Figure 1 is difficult to follow. It is unclear how the means are being calculated. For example, it seems like in examples 2 and 3, the means between conditions should be different. I was also confused in example 4 because the difference in probability between the two conditions in time bin 2 seems tiny. I was also confused because it still looked like there was a big difference in probability between the two conditions at bin 5. Was bin 6 by chance what was meant instead of bin 5? It also still looked like there were differences later on such as in bin 7. Example 5 also left me confused. It looked to me like the effect started at bin 4 and ended at bin 8. Clarification (and perhaps simplifying) this figure would be useful.

Response:

> The description for figure 2 is unclear, particularly with respect to the contrasting knowledge that can be gained from the survival function and hazard function

Response:

### 2025-06-14 manuscript/ResponseReview1.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{framed}
\usepackage{xcolor}
\let\oldquote\quote
\let\endoldquote\endquote
\renewenvironment{quote}{\begin{leftbar}}{\end{leftbar}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Review 1: Author responses},
  pdfauthor={Panis \& Ramsey},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Review 1: Author responses}
\author{Panis \& Ramsey}
\date{12.06.2025}

\begin{document}
\maketitle

\section{Editor}\label{editor}

\begin{quote}
15-Mar-2025

Dear Dr.~Panis:

Thank you for submitting your Tutorial (AMPPS-24-0232) entitled ``Event
History Analysis for psychological time-to-event data: A tutorial in R
with examples in Bayesian and frequentist workflows'' to Advances in
Methods and Practices in Psychological Science (AMPPS). I have now
received input from three experts in the field, and you will find the
reviews below this message and attached. As you will see, all three
reviewers believe your paper can make a strong contribution to the
literature. Prior to considering the reviewers' input, I read your paper
to form an independent opinion. I share the positive sentiments about
this work and invite you to revise and resubmit your manuscript for
further consideration at AMPPS.
\end{quote}

\begin{quote}
The reviews are through and thoughtful, and I will not reiterate all the
details here. For my part, I would like your revision to focus on
accessibility. Your paper is a bit challenging-- it's long and a little
tough to get through. What can make it easier for less familiar readers?
Perhaps more of a basic set up earlier? A glossary defining specific
terms (e.g., what is an event? how is it defines? censoring? hazards
vs.~survival curves, conditional probabilities, etc.)? Also, clarify
when readers need to shift from frequentist to Bayesian model\ldots{}
and why? The main consideration here is that wish to make the paper much
more accessible-- from start to finish-- for readers who are novices and
entirely unfamiliar with this area. This can be done any number of ways,
and R3 also includes some suggestions for making these reviews. I
encourage you to have the revised paper read and pre-reviewed by
colleagues who are novices in this area with the explicit goal of making
the paper accessible to them-- this will help you hit the more for our
broad readership.
\end{quote}

\begin{quote}
If you choose to submit a revision, please include include a letter
detailing your point-by-point responses to each reviewer comment and
indicating how you changed the manuscript to address them. The revised
manuscript may undergo further peer review. We ask that you submit your
revision within three months. Please let us know if you will not be able
to meet this deadline.
\end{quote}

\begin{quote}
To submit your revision, log into
\url{http://mc.manuscriptcentral.com/ampps} and enter your Author
Center, where you will find your manuscript title listed under
``Manuscripts with Decisions.'' Under ``Actions,'' click on ``Create a
Revision.'' Your manuscript number has been appended to denote a
revision.
\end{quote}

\begin{quote}
IMPORTANT: Your original files are available to you when you upload your
revised manuscript. Please delete any redundant or outdated files before
completing the submission.
\end{quote}

\begin{quote}
Once again, thank you for submitting your manuscript to AMPPS and I look
forward to receiving your revision.
\end{quote}

\begin{quote}
Sincerely,
\end{quote}

\begin{quote}
David A. Sbarra, Ph.D. Editor, Advances in Methods and Practices in
Psychological Science
\end{quote}

@@

We tried to focus on accessibility in the revision of our Tutorial
AMPPS-24-0232.

\section{Reviewer: 1}\label{reviewer-1}

\begin{quote}
Comments to the Author It was a pleasure to review this manuscript as
someone who has spent years applying survival analysis methods to
psychological data. This is an ambitious manuscript and I believe it
achieves its goals. The key contributions of this manuscript from my
perspective are the creation of custom functions for data wrangling, and
the section on planning studies/simulations for power analysis (of
course, in addition to the introduction to this category of methods and
the practical tutorials provided. I have some comments on this work that
I believe could strengthen this manuscript's contribution.
\end{quote}

\begin{quote}
General comments:
\end{quote}

\begin{quote}
The Introduction is really nicely framed in terms of providing
compelling information for why this class of statistical models would be
useful in some specific areas of psychology (i.e., reaction time data).
Given that there are already a couple of tutorial-types of articles
introducing specific applications of survival analysis in developmental
psychology (references below, both regarding the use of these methods
for behavioural observations of emotional expressions in a multilevel
framework for recurring events), it would be helpful to further frame
the unique contribution that this work adds. The authors may also
consider further emphasizing the specific theoretical contributions that
using these methods may lead to in the areas of reaction time data etc.
\end{quote}

@@

Response: In the revised ms. We added the following section: In this
tutorial we provide code to implement discrete-time hazard models in a
Bayesian framework, with a focus on cognitive experimental psychology.

\begin{quote}
Lougheed, J. P., Benson, L., Ram, N., \& Cole, P. M. (2019). Multilevel
survival analysis: Studying the timing of children's recurring
behaviors. Developmental Psychology, 55(1), 53--65.
\url{https://doi.org/10.1037/dev0000619}
\end{quote}

\begin{quote}
Stoolmiller, M. (2016). An introduction to using multivariate multilevel
survival analysis to study coercive family process. In T. J. Dishion \&
J. J. Snyder (Eds.), The Oxford handbook of coercive relationship
dynamics (pp.~363--378). Oxford University Press.
\end{quote}

\begin{quote}
Stoolmiller, M., \& Snyder, J. (2006). Modeling heterogeneity in social
interaction processes using multilevel survival analysis. Psychological
Methods, 11(2), 164--177.
\url{https://doi.org/10.1037/1082-989X.11.2.164}
\end{quote}

@@

Response: We added these references to the ms.

\begin{quote}
Similarly, it would be helpful to make it clear that this manuscript
pertains to estimating single-event occurrences rather than recurring
events (which are described in the Lougheed and Stoolmiller tutorials
mentioned above)
\end{quote}

@@

Response: On page x of the revised ms., we mention that we estimate
single-event occurrences that are repeatedly measured in each
participant.

\begin{quote}
Specific comments:
\end{quote}

\begin{quote}
The authors use several different terms to refer to time-to-event models
throughout the manuscript. The title refers to EHA, and the running head
refers to Hazard analysis. Survival analysis is also a common term in
Psychology (see tutorials by Lougheed et al., 2019, and Stoolmiller \&
Snyder, 2006). Disambiguating the various terms early in the manuscript
and then constantly using one term throughout the manuscript would be
helpful.
\end{quote}

@@

Response: We disambiguate the various term early in the revised ms.
(page x).

\begin{quote}
Page 5: it would be helpful to refer to the same time unit in reference
to Figure 1, which shows time bins along the x-axis but the examples
discussed in the manuscript refer to units of milliseconds.
\end{quote}

@@

Response: We refer to the same time unit in Figure 1 and the text.

\begin{quote}
Page 7: The authors state that they are not aware of EHA tutorials for
psychological time to event data. Previoulsy mentioned tutorial articles
by Lougheed and Stoolmiller could be mentioned here--- these are
demonstrating applications to behavioural observation data within
psychology.
\end{quote}

@@

Response: We added these 3 references to the revised ms.

\begin{quote}
Page 8: I'm curious about the decision to focus on discrete time rather
than continuous time models. It would help the reader to explain the
decision either way, or refer the reader to Lougheed et al., 2019 who
discuss this distinction and choices that researchers may make regarding
how to handle decisions around time given data formats. I don't work
with reaction time data, but it seems to me that many experiments in
this area would collect data in continuous time intervals, and it's not
clear to me what the benefit of discretizing into time bins is. Now that
I've read the entire manuscript, I see that some information is
presented on this on Page 50, but it would be helpful to make this
information known to the reader earlier in the manuscript.
\end{quote}

@@

Response: We address the issue of treating time as discrete or
continuous on page x of the revised ms.

\begin{quote}
Page 8-9, excellent and accessible explanations of the survival function
in the discrete time framework.
\end{quote}

@@

Response:

\begin{quote}
Page 11: In my experience, researchers in psychology are generally not
familiar with the concept of right censored data. This term is first
used on page 11 with no explanation--- a brief, even parenthetical,
definition would be helpful. For example, the parenthetical definition
on page 48 could be provided on page 11.
\end{quote}

@@

Response: We define right-censoring on page x of the revised ms.

\begin{quote}
Taken together, this is an excellent manuscript with the potential to
further bring survival analysis methods into Psychology, where
time-to-event and right-censored cases abound, yet very few people seem
to consider these issues.
\end{quote}

\section{Reviewer: 2}\label{reviewer-2}

\begin{quote}
Comments to the Author Please see attached file for comments.
\end{quote}

\begin{quote}
Reviewer report for AMPPS-24-0232: ``Event History Analysis for
psychological time-to- event data: A tutorial in R with examples in
Bayesian and frequentist workflows'' Summary: The authors present a
tutorial paper on the analysis of discrete time-to-event outcomes. They
motivate their tutorials with a description of psychological experiments
that produce time-to-event data and explain how simply summarizing
event-time outcomes using means results in a loss of potentially
important information. They then explain how more information can be
gleaned from event-time data if they are summarized using event history
analysis methods. They focus specifically on the analysis of discrete
time-to-event outcomes. They provide code for conducting the analyses in
the tutorials on their Github. Overall, I like the idea of this paper
and strongly agree that using methods appropriate for analyzing
time-to-event data, rather than simply summarizing event-time outcomes
using means, is very important and currently underused. While I like the
premise of the paper and think that such a paper could be a valuable
contribution to the literature, I do think the paper has room for
improvement in a number of areas (see comments below). The writing is
approachable (which is advantageous for a tutorial paper) but some text
in the paper was repetitive and could be a bit more polished; additional
editing with a focus on clarity and brevity would be beneficial.
\end{quote}

@@

Response: We removed repetitive text, and try to focus on clarity and
brevity in the revised ms.

\begin{quote}
Major comments: 1. Some discussion on the choice of bin width is needed.
Since all of the example time-to-event data is collected in continuous
time, the analyst must discretize it by defining bins. Defining bins is
a key step in the analysis pipeline and so should not be overlooked. It
seems like results---and potentially conclusions drawn---may depend on
the choice of bin width. What factors should the analyst consider when
defining the width of the bins? Does the number of bins matter? Or is
the number of events per bin important? What about other factors? Should
analysts (or potentially these authors too) consider multiple bin widths
when conducting an analysis? An example comparing results when bin width
varies could be both interesting and informative.
\end{quote}

@@

Response: We discuss the issue of defining bins on page x of the revised
ms.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  In the discussion, the authors mention that they choose to present a
  tutorial on discrete time analysis methods because they believe these
  are easier to understand. My personal impression is that awareness of
  continuous time analysis methods is slightly higher and so researchers
  may be more familiar with methods such as Cox models, even if they
  have not used them themselves; however, this could vary by specific
  sub-field. I agree with the authors that the interpretation of the
  hazard is easier in the discrete time setting than in the continuous
  time setting, but I do think that the distinction between the
  interpretation of the hazard in continuous vs discrete time should be
  discussed in the introduction. When I first read the introduction, I
  was confused as to why the focus was on discrete time methods, when
  the outcome measures collected in the motivating data were collected
  in continuous time. I would suggest either of the following: (a)
  provide concrete examples of when truly discrete timeoutcomes are
  collected to strengthen the motivation of the paper and justification
  for use of discrete time analysis methods or (b) keep the motivating
  examples as- is and include a short comparison of discrete
  vs.~continuous time analysis methods to help explain your reasoning
  for choosing to focus on discrete time analysis methods.
\end{enumerate}
\end{quote}

@@

Response: We include a short comparison of both methods on page x of the
revised ms.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Censoring is an important part of time-to-event data and so more
  description of censoring---and how it is (not) accounted for in
  ``orthodox'' analysis methods vs. discrete time-to-event analysis
  methods would be useful, beyond the already- included mention of
  censoring in Section 5.3.
\end{enumerate}
\end{quote}

@@

Response: We define (right-)censoring on page x of the revised ms.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Section 1.2 is a bit long. Please shorten this section to improve the
  clarity of the text and reduce redundancy. There is also overlap
  between text in Section 1.2 and the start of Section 4.
\end{enumerate}
\end{quote}

@@

Response: We shortened this section and removed the overlap.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Section 2.3.2. I like the idea of making a connection with
  experimental design, but am having trouble seeing how the discussion
  in this section is specific to EHA and how these comments do not apply
  also to the ``orthodox method''. I think this section could be
  improved by providing more specific distinctions between experimental
  design implications for EHA vs.~the ``orthodox method''.
\end{enumerate}
\end{quote}

Response: We added the following section on page x of the revised ms.:
\ldots{}

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
  Line 517: What are model weights? Please explain briefly for the
  reader, including how the interpretation differs for loo and waic.
  Also, in the provided reference (Kruz 2023a), I was only able to find
  a discussion of dWAIC, rather than WAIC, so could you please include a
  definition of or reference for dWAIC in your context? Lastly, when
  reporting the model weights in the example in the text, I'd suggest
  including more than 2 digits so that it is clear if the reported
  weights are actually 0 and 1 or if rounding has just made them appear
  so.
\end{enumerate}
\end{quote}

@@

Response: We explain model weights on page x, include a reference for
dWAIC, and report the weights with 3 digits.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  Tutorials 3a and 3b: Rather than simply stating that the frequentist
  models did not converge and resulted in singular fits, it would be
  helpful to provide more discussion. Do you have an idea why the models
  did not converge? Since this is a tutorial paper, more explanation is
  needed. These frequentist examples don't add much to the paper in
  their current form, so I'd either suggest either (a) providing some
  suggestions to the reader of alternative approaches or specific
  strategies that could be used to define a model that will converge or
  (b) removing these examples from the main text and directing the
  reader to the supplement instead.
\end{enumerate}
\end{quote}

@@

Response: We removed the frequentist examples from the main text into
the supplementary material.

\begin{quote}
Minor comments: 1. Is ``mean-average comparison'' standard terminology?
This phrase seems redundant to me, since means and averages are the same
thing.
\end{quote}

@@

Response: We removed this term.

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Figure 1: In the caption, please state what the small error bars in
  the inset plots are showing.
\end{enumerate}
\end{quote}

@@

Response:

\begin{quote}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Please consider updating all plots so that they are understandable
  when viewed in black and white. In addition to using different colors,
  I'd suggest distinguishing the conditions using shapes (e.g., circles
  vs.~triangles) and different line type (e.g., solid vs.~dashed lines).
\end{enumerate}
\end{quote}

@@

Response:
